A game-theory modeling approach to utility and strength of interactions dynamics in biomedical research social networks

### 2017-04-05 casm.tex
Finished conclusions and added Conte et al reference

Another interesting behavior explored by game-theoretical modeling and
complex networks has been done by Li and Cao \cite{Li-etal:2009}. This
authors analyze the role of punishment in a science collaboration
network. More precisely, they implemented an ultimatum bargaining
model to understand the impact of the topology of a social network and
the emergence of a population of punishers. In the ultimatum game,
there are two players, one of them (the proponent) suggests to split a
sum in two and its connected neighbor (the respondent) can decide to
accept or reject the offer as it may be perceived more or less
fair. If the respondent accepts, the sum is divided according to the
offer. If the offer is rejected both players get nothing. According to
the authors, the act of rejecting a selfish offer can be interpreted
as a punishment since the player that rejects it is paying the cost of
not receiving anything, but because of a selfish offer the proponent
looses even more. Their results show that a phase transition will
emerge in the population of punishers as the degree of selfishness
increses among players. Moreover, they found a dependence relation
between the tolerance to selfish behaviors and the topologies of
networks.}\\

@@

{\color{red}In a similar vein to the work of Li and Cao,} we explore the network
effect on the distribution of players having access to certain amount of
resources shared by other players in the network and the distribution of the
strength of interactions among them. Particularly, we implemented two games
played simultaneously: one for maximizing individual utility based on the
iterated Prisoner's Dilemma (PD); the other, a coordination game for maximizing the
connection strength between players. We are interested in how they affect each
other in the context of a network of scientific collaboration under the idea
that while researchers are interested in maximizing their utilities, they also
know that it is important to invest in building collaborative
relationships. \\ %This is important because what we have seen

@@

  On these grounds, part of our future work is based on the results
  presented here. We would like to identify researchers in our
  simulations and corroborate their situation in the model and in the
  real world. We are also interested in going back to the field and
  interviewing those groups with an interesting behavior found in our
  simulations, we would probably follow a similar strategy as the one
  developed in \cite{Hara:2003}. Finding communities beyond the level
  of the groups is an important task. We think that there are many
  possibilities that emerge from the integration of different
  methodologies. Moreover, studying social processes in science is
  particularly attractive due to the amounts of data already available
  that can be easily collected. This is a privilege because
  simulations can be designed on real world data, something that only
  very recently has become possible \cite{Barabasi:2012}.\\

@@

No doubt social sciences are becoming more ``computational''. This can
be seen everywhere, but curiously enough, disciplines like physics
and computer science are moving towards the social
sciences and not so much the other way around. It might be the case
that the pioneering disciplines in the computational social sciences
will set the agenda, an agenda that will apparently be mostly based on
taking advantage of big data and on hypothesis-free approaches. We
believe that the social sciences have important questions that should
be added to that agenda and those questions may not be answered only
by big data techniques but they may require creating models and
simulations in the style of the best hypothesis-driven research.

@@

We acknowledge the help of Ma. del Pilar Galarza and Francisco Allende
for their work curating databases of researchers. We are grateful to
Be Luecke for copy editing the manuscript.

>>>>

Another interesting behavior explored by game-theoretical modeling and
complex networks has been done by Li and Cao \cite{Li-etal:2009}. These
authors analyze the role of punishment in a science collaboration
network. More precisely, they implemented an ultimatum bargaining
model to understand the impact of the topology of a social network and
the emergence of a population of punishers. In the ultimatum game,
there are two players, one of them (the proponent) suggests to split a
sum in two and its connected neighbor (the respondent) can decide to
accept or reject the offer as it may be perceived more or less
fair. If the respondent accepts, the sum is divided according to the
offer. If the offer is rejected both players get nothing. According to
the authors, the act of rejecting a selfish offer can be interpreted
as a punishment since the player that rejects it is paying the cost of
not receiving anything, but because of a selfish offer the proponent
looses even more. Their results show that a phase transition will
emerge in the population of punishers as the degree of selfishness
increses among players. Moreover, they found a dependence relation
between the tolerance to selfish behaviors and the topologies of
networks.}\\

@@

{\color{red} In our model we explore the network effect on two issues: a) the distribution of the amount of
resources players can access from other players and, b) the distribution of the strength of
interactions among players as they interact. Also, in a similar
vein to the work of Li and Cao, we explore how resistant are the different
topologies to defection as it becames harder for the population to cooperate.}
Particularly, we implemented two games played simultaneously: one for maximizing
individual utility based on the iterated Prisoner's Dilemma (PD); the other, a
coordination game for maximizing the connection strength between players. We are
interested in how they affect each other in the context of a network of
scientific collaboration under the idea that while researchers are interested in
maximizing their utilities, they also know that it is important to invest in
building collaborative relationships.
\\ %This is important because what we have seen

@@

@@

Our future work is based on the results presented here. We would like to
identify researchers in our simulations and corroborate their situation in the
model and in the real world. We are also interested in going back to the field
and interviewing those groups with an interesting behavior found in our
simulations, we would probably follow a similar strategy as the one developed in
\cite{Hara:2003}. Another idea we would like to follow is related to the
parameter $\nu$. As it is now implemented in the model, $\nu$ is an external
parameter to the system but in future work we would like to explore the
possibility of $\nu$ being controlled by the dynamics of the system. The
assumption behind this idea is that society self-imposes norms and it
dynamically regulates how difficult is for its members to comply. In our model
we used this parameter to control how hard or easy is for a player to
cooperate. Finding communities beyond the level of the groups is an important
task. We think that there are many possibilities that emerge from the
integration of different methodologies. Moreover, studying social processes in
science is particularly attractive due to the amounts of data already available
that can be easily collected. This is a privilege because simulations can be
designed on real world data, something that only very recently has become
possible \cite{Barabasi:2012}.\\

No doubt social sciences are becoming more ``computational''
\cite{Conte-etal:2012}. This can be seen everywhere, but curiously enough,
disciplines like physics and computer science are moving towards the social
sciences and not so much the other way around. It might be the case that the
pioneering disciplines in the computational social sciences will set the agenda,
an agenda that will apparently be mostly based on taking advantage of big data
and on hypothesis-free approaches. We believe that the social sciences have
important questions that should be added to that agenda and those questions may
not be answered only by big data techniques but they may require creating models
and simulations in the style of the best hypothesis-driven research.

@@

We acknowledge the help of Ma. del Pilar Galarza and Francisco Allende
for their work curating databases of researchers. We are grateful to
Be Luecke for copy editing the manuscript. We would also like to thank to the
reviewers for their comments and suggestions.
